NumPy---Array Programming in Python

### 2020-01-13 body.tex
Add new section title
Start expanding Supporting the Ecosystem

{\em Strides} are necessary to interpret computer memory, which stores elements
linearly, as $n$-dimensional arrays.
It describes the number of bytes to move forward in memory to jump from row to
row, column to column, and so forth.
Consider, for example, a 2-D array of floating point numbers with shape
$(4, 3)$, where each element occupies 8 bytes in memory.
To move between consecutive columns we need to jump forward 8 bytes in memory,
and to access the next row $3 \times 8 = 24$ bytes.
The strides of that array are therefore $(24, 8)$.

@@

%Once data is viewed as an $n$-dimensional array, more abstract operations such
%as reductions can be defined.

%(e.g., summation across columns, see Fig.~\ref{fig:reductions}) products, and %so forth, are
%defined.

@@

Users predominantly interact with NumPy arrays using {\em indexing} (to access
sub-arrays or individual elements), {\em operators} (e.g., $+$, $-$, $\times$,
and, for matrix multiplication, $@$), as well as {\em utility functions};
together, these provide an easily readable, expressive, high-level API for
array programming.
Exposing these array programming primitives in Python---an interpreted
language---creates an ideal environment for exploratory data analysis where
users may inspect, manipulate, and visualize their data
\cite{perez2007ipython}.

@@

@@

>>>>

{\em Strides} are necessary to interpret computer memory, which stores elements
linearly, as $n$-dimensional arrays.
It describes the number of bytes to move forward in memory to jump from row to
row, column to column, and so forth.
Consider, for example, a 2-D array of floating point numbers with shape
$(4, 3)$, where each element occupies 8 bytes in memory.
To move between consecutive columns we need to jump forward 8 bytes in memory,
and to access the next row $3 \times 8 = 24$ bytes.
The strides of that array are therefore $(24, 8)$.  NumPy is able to
store arrays in both so-called ``C'' and ``Fortran''-order, iterating
first over either rows or columns.  This allows external libraries
written in those languages to directly access NumPy array data in memory.

@@

@@

\section*{Supporting the ecosystem}

Users predominantly interact with NumPy arrays using {\em indexing} (to access
sub-arrays or individual elements), {\em operators} (e.g., $+$, $-$, $\times$,
and, for matrix multiplication, $@$), as well as {\em utility functions};
together, these provide an easily readable, expressive, high-level API for
array programming.

Various libraries build on and extend this functionality to further enrich the ecosystem of available tools.  For example, SciPy adds various mathematical, scientific, and engineering routines; Pandas provides indexed arrays, known as dataframes; xarray implements arrays with labeled axes; and Matplotlib generates publication-ready figures and visualizations.  Employing these low-level utilities are several widely used technique-specific libraries, including scikit-learn (machine learning), scikit-image (image processing), statmodels (statistics),  networkx (graphs and complex networks), and Napari (volumetric image visualization) \fixme{citations for all these}.

Finally, domain-specific libraries and applications are built on this diverse foundation of tooling; these include astropy (astronomy) \fixme{citation}, NiPy (neuroimaging) \fixme{citation}, and pangeo (earth science) \fixme{citation}.

% ecosystem
% core: scipy, pandas, matplolib ;
% scikits: scikit-learn, scikit-image, statmodels, networkx
% domain specific: astropy (2 citations?), neuroscience (3 citations?), pangeo, napari, etc.
% Ipython / Jupyter

\fixme{Maybe some wording around the actual *support* element?}

Exposing array programming primitives, as well as the surrounding ecosystem of tools, in Python---an interpreted language---creates an ideal environment for interactive, exploratory data analysis where users may iteratively inspect, manipulate, and visualize their data \cite{perez2007ipython}.

@@

% Fortran / C
% IDL / FITS / ROOT / Parquet / Arrow ?

@@

% teaching (neuroimaging paper / etc.?)
